---
title: "Minimum values of the Rooted Quartet index"
output: html_document
---

The function `Max.rQI.Rec` applied to any positive natural number `n` 
computes the vector of values $rQI(B_k)$, for $k=1,...,n$
by means of the recurrence
$$
rQI(B_n) = rQI(B_{\lceil n/2\rceil}) + rQI(B_{\lfloor n/2\rfloor}) + \binom{\lceil n/2\rceil}{2}\binom{\lfloor n/2\rfloor}{2}
$$
```{r}
Max.rQI.Rec=function(n){
  x=rep(0,n)
  for (i in 2:n){
    x[i]=x[floor(i/2)]+x[ceiling(i/2)]+choose(floor(i/2),2)*
      choose(ceiling(i/2),2)
  }
  x}
```

For instance, the first 1000 values of $rQI(B_k)$ according to this recurrence are:
```{r}
Max.rQI.Rec(1000)
```

The function `Max.rQI` applied to any positive natural number `n`
computes the value $rQI(B_n)$ using its closed formula given in our paper "The total cophenetic and rooted quartet indices of the maximally balanced trees" (Theorem 4).

```{r}
Max.rQI=function(n){
  BD=floor(log2(n))+1
  VV=as.integer(intToBits(n))[1:BD]
  M=(0:(BD-1))[VV==1]
  s=length(M)
  MM=floor(n/2^(M+1))
  F0=function(i){2^(2*M[i]-1)*(2^M[i]+3)}
  F1=function(i){(10*n+2^(3*M[i]+1)+2^(2*M[i]+2)-12*2^(M[i]+1))*MM[i]}
  F2=function(i){(3*n+2^(3*M[i]+1)-15*2^M[i])*MM[i]^2}
   F3=function(i){2^(M[i]+1)*MM[i]^3}
  rQIB=(1/504)*n*(9*n^3 -46*n^2+51*n+48*s-78)+1/63*sum(sapply(1:length(M),FUN=F0))+1/42*sum(sapply(1:length(M),FUN=F1))+1/21*sum(sapply(1:length(M),FUN=F2))-1/7*sum(sapply(1:length(M),FUN=F3))
  rQIB}
```

For instance, the first 1000 values of $rQI(B_k)$ computed with our closed expression are
```{r}
sapply(1:1000,FUN=Max.rQI)
```

Unfortunately, in this case we cannot check directly with the `==` function the equality of both sequences, because of floating point calculations in our definition of `Max.rQI`:

```{r}
Max.rQI.Rec(10)
sapply(1:10,FUN=Max.rQI)
Max.rQI.Rec(10)==sapply(1:10,FUN=Max.rQI)
```

But we can check their equality by first rounding to integers the values of the second sequence (and noting that they are actually integer numbers, up to floating point approximations).

```{r}
Max.rQI.Rec(10)
sapply(1:10,FUN=Max.rQI)
Max.rQI.Rec(10)
round(sapply(1:10,FUN=Max.rQI))
Max.rQI.Rec(10)-round(sapply(1:10,FUN=Max.rQI))
max(Max.rQI.Rec(1000)-round(sapply(1:1000,FUN=Max.rQI)))
```




Their log-log plot is
```{r,out.width="90%",scale=2}
plot(4:1000,Max.rQI.Rec(1000)[-(1:3)],pch=20,cex=0.2,xlab="n",
     ylab=expression(rQI(B[n])),main="Maximum values of the Rooted Quartet index",log="xy")
```
