---
title: "Minimum values of the Total Cophenetic index"
output: html_document
---

The function `Min.Phi.Rec` applied to any positive natural number `n` 
computes the vector of values $\Phi(B_k)$, for $k=1,...,n$
by means of the recurrence
$$
\Phi(B_n)= \Phi(B_{\left\lceil{n}/{2}\right\rceil}) +\Phi(B_{\left\lfloor{n}/{2}\right\rfloor})+\binom{\left\lceil{n}/{2}\right\rceil}{2}+\binom{\left\lfloor{n}/{2}\right\rfloor}{2}
$$

```{r}
Min.Phi.Rec=function(n){
x=rep(0,n)
for (i in 2:n){
x[i]=x[floor(i/2)]+x[ceiling(i/2)]+choose(floor(i/2),2)+
  choose(ceiling(i/2),2)
}
x}
```

For instance, the first 1000 values of $\Phi(B_k)$ according to this recurrence are:
```{r}
Min.Phi.Rec(1000)
```

The function `Min.Ph` applied to any positive natural number `n`
computes the value $\Phi(B_n)$ using its closed formula given in our paper "The total cophenetic and rooted quartet indices of the maximally balanced trees" (Theorem 3).

```{r}
Min.Phi=function(n){
  BD=floor(log2(n))+1
  VV=as.integer(intToBits(n))[1:BD]
  M=(0:(BD-1))[VV==1]
  F=function(i){2^(M[i]-1)*(M[i]+2*(length(M)-i))}
  PhiB=choose(n,2)-sum(sapply(1:length(M),FUN=F))
  PhiB}
```

For instance, $\Phi(B_{10^6})$ is

```{r}
Min.Phi(10^6)
```
and the first 1000 values of $\Phi(B_k)$ computed with our closed expression are
```{r}
sapply(1:1000,FUN=Min.Phi)
```

Both sequence are equal:
```{r}
as.logical(min(sapply(1:1000,FUN=Min.Phi)==Min.Phi.Rec(1000)))
```


Their log-log plot is
```{r,out.width="90%",scale=2}
plot(3:1000,Min.Phi.Rec(1000)[-(1:2)],pch=20,cex=0.2,xlab="n",
     ylab=expression(Phi(B[n])),main="Minimum values of the Total Cophenetic index",
     log="xy")
```
